---
title: "Using Rev Knitr"
author: "April Wright"
date: "4/22/2021"
output: html_document
---

```{r setup, include=TRUE}
library(RevR)
knitr::opts_chunk$set(echo = TRUE, eval=TRUE, echo=TRUE)
InitRev("/Users/april/software/rb")
KnitRev()
```

## RevBayes and R

Calling external software from R can be a bit tricky, and RevBayes is no exception. In order to successfully use RevBayes in R, you must have RevBayes [installed](https://revbayes.github.io/download) on your computer. You must also know the system path to the RevBayes executeable. For example, on my computer, my RevBayes executeable is called "rb" (this will be the case for Mac and Linux users), and it is stored in my software folder. If you are on PC, your copy will be called "rb.exe."

In the above chunk, you will see the command `InitRev`. This is a function in the `RevKnitR` R package. This R package can be installed using the popular `devtools` R package like so:

```{r, include=FALSE}
devtools::install_github("Paleantology/RevR")
```

The `InitRev` function creates a RevBayes running environment, which will allow you to interact with RevBayes through R. `InitRev` takes one argument: where RevBayes lives on your computer. Delete my sample path and enter your path above.

The next line, `KnitRev` establishes a KnitR environment to render Rev code in the attractive KnitR format many of us are used to.

## Using RevBayes in a KnitR Chunk

RevBayes can be used in a KnitR chunk by changing the header to `rb` instead of `r`. In the below chunk, we create an object called `example` and use the assignment operator to give it the value 1. Then we print it. 
```{rb}
example <- 1.0
example
```
This is not an overtly useful thing to do, however. Let's erase the previous chunk using the `ClearRev()` function. This removes prior code and objects from the RevBayes environment. Very handy if you make a mistake!

```{r}
ClearRev()
```

We could, instead, choose to do something a little more useful. How about reading in a data matrix and making a quick starting tree? 

```{rb}
morpho <- readDiscreteCharacterData("../data/Cinctans.nex")
num_taxa <- morpho.size()
num_branches <- 2 * num_taxa - 2
taxa <- morpho.names()
br_len_lambda ~ dnExp(0.2)
phylogeny ~ dnUniformTopologyBranchLength(taxa, branchLengthDistribution=dnExponential(br_len_lambda))
phylogeny
```
Anything entered in an `rb` block will be interepreted as Rev code, and all the normal Rev syntax will apply. For a nice overview of Rev language and syntax, please see [this tutorial](https://revbayes.github.io/tutorials/intro/rev). 

One thing researchers are often interested in doing is making an object in Rev and then viewing it in R. 
```{r}
RevR::doRev('morpho <- readDiscreteCharacterData("../data/Cinctans.nex");
num_taxa <- morpho.size();
num_branches <- 2 * num_taxa - 2;
taxa <- morpho.names();
br_len_lambda ~ dnExp(0.2);
phylogeny ~ dnUniformTopologyBranchLength(taxa, branchLengthDistribution=dnExponential(br_len_lambda));
phylogeny', coerce=TRUE)
library(phytools)
phylo <- RevEnv$phylogeny
phylo
typeof(phylo)
#plotTree(phylogeny)
```
